Shared Subspace Model

### 2016-05-19 paper/mscov.tex
Updated paper

\section{Inference}
\subsection{An Expectation-Maximization Algorithm}
\label{sec:em}

We start by outlining a computationally efficient algorithm for
estimating the shared subspace, $VV^T$, noise $\sigma_k^2$ and
projected covariance matrix $\psi_k$.

First, note that the full likelihood can be written as

\begin{align}
p(S_1, ... S_k | \Sigma_k,n_k) &\propto \prod_{k=1}^K |\Sigma_k|^{-n_k/2}etr(-\Sigma_k^{-1}\mathbf{S}_k/2)  \\
&\propto \prod_{k=1}^K  |\Sigma_k|^{-n_k/2}etr(-(\sigma_k^2(V\psi_kV^T +
  I))^{-1}\mathbf{S}_k/2) \\
&\propto \prod_{k=1}^K  |\Sigma_k|^{-n_k/2}etr(-\left[V(\psi_k +
  I)^{-1}/\sigma_k^2 V^T + (I-VV^T)/\sigma^2_k\right]\mathbf{S}_k/2)
  \\
&\propto \prod_{k=1}^K  (\sigma_k^2)^{-n_k(p-r)/2}|M_k|^{-n_k/2}etr(-\left[VM_k^{-1}V^T + \frac{1}{\sigma^2_k} (I-VV^T)\right]\mathbf{S}_k/2) 
\end{align}

@@

% \begin{align}
% p(V|...) &\propto etr(\sum_{k=1}^K[O_k\Omega_kO^T_k]V^TS_k/(2\sigma^2_k)]V)\\
% \end{align}

We propose a computationally efficient expectation-maximizatino (EM)
algorithm, to estimate the relevant quantities.  In particular we
maximize the marginal posterior distribution of $V$, treating the
$M_k^{-1}$ and $\frac{1}{\sigma_k^2}$ as the ``missing'' parameters to
be imputed.  We assume independent Jeffreys' prior distributions for
both $\frac{1}{\sigma_k^2}$ and $M_k^{-1}$ and a uniform prior for $V
\in \mathcal{V}_{s, p}$.  In particular we let
$p(\frac{1}{\sigma_k^2}) \propto \sigma_k^2$ and
$p(M_k^{-1}) \propto |M_k^{-1}|^{-(r+1)/2}$.  From the likelihood it
can easily be shown that the conditional posterior for $M_k^{-1}$ is

$$p(M_k^{-1} | V) \propto |M_k^{-1}|^{(n_k - r -1)/2}etr(-1/2(M_k^{-1}V^T\mathbf{S}_kV)) $$

@@

$$p\left(\frac{1}{\sigma^2} | V\right) \propto \left(\frac{1}{\sigma_k^2}\right)^{n_k(p-r)/2-1}etr\left(-\frac{1}{2\sigma^2_k} (I-VV^T)\mathbf{S}_k\right)  $$

@@

% \subsection{Alternative formulation}
% $M_k = \sigma_k^2(\Phi_k + I)$ where 
% $$ \Phi_k =\left( \begin{array}{cc}
% \Psi_k & 0  \\
% 0 & D  \end{array} \right)$$

% Letting $V = [V_1\, V_2]$

% \begin{eqnarray*}
% p(M_k^{-1} | V) &\propto \prod_k |M_k^{-1}|^{n_k/2}etr(-1/2(M_k^{-1}V^T\mathbf{S}_kV)) \\
%                 & = (|\psi_k + I|^{-1}|D +
%                   I|^{-1})^{n_k/2}etr(-1/2((\psi_k+I)^{-1}V_1^T\mathbf{S}_kV_1
%                   -1/2(D+I)^{-1}V_2^T\mathbf{S}_kV_2)) \\
% \end{eqnarray*}

% $$p(\frac{1/d_i} | V) \propto \prod_k (d_i + 1)^{-n_k/2}exp\left(\frac{1}{-1/2(d_i+1)}(V_2^T\mathbf{S}_kV_2)_{i,i}\right) $$

\subsection{Bayesian Inference}
\label{sec:bayes}

Although the EM algorithm presented in the previous section yields
point estimates of $\psi_k$, it does not lead to natural uncertainty
bounds for these estimates.  To quantify the uncertainty, we consider Bayesian
inference for $\psi_k$, given the
subspace $VV^T$.  Note that when the subspace is known, the posterior
distribution of $\Sigma_k$ is conditionally independent from the other
groups, i.e.  $P(\Sigma_{k} | V, \bl S_k, \Sigma_{-k}) = P(\Sigma_{k} | V, \bl S_k)$.
Thus, we can independently estimate the conditional posterior
distribution for each group.  In this paper, we focus on models based
on the eigendecomposition of $\psi_k$.  In this setting, each covariance
$\Sigma_k$ is a spiked covariance matrix where Equation \ref{eqn:sspsi}
can be written as

@@

Here, $\Lambda_k$ is an $r \times r$ diagonal matrix of eigenvalues,
and $O_k \in \mathcal{V}_{s,r}$ is the matrix of eigenvectors of
$\psi_k$.  For any individual group, this corresponds to the original
spiked covariance model (Equation \ref{eqn:spiked}) with $U_k = VO_k$.

@@

\begin{equation}
p(S | \Sigma_k,n_k) \propto |\Sigma_k|^{-n/2}etr(-\Sigma_k^{-1}\mathbf{S}_k/2)
\end{equation}

@@

Where the diagonal matrix $\Omega = \Lambda(I+\Lambda)^{-1}$, e.g. $\omega_i = \frac{\lambda_i}{\lambda_{i}+1}$.

 Further,

@@

where the second line is due to Sylvester's determinant theorem.

Substituting for $\Sigma^{-1}_k$ and $|\Sigma_k|$ we have the likelihood of $V$, $O_k$, $\Lambda_k$ and $\sigma_k^2$

We can write the likelihood of $V$, $O_k$, $\Lambda_k$ and
$\sigma_k^2$ given $Y_k$ using Equation \ref{eqn:rplik} by replacing
$U_k$ with $VO_k$: 
\begin{equation}
 L(\sigma_k^2,\bl V , O_k \Omega_k : \bl Y_k) \propto
    (\sigma_k^2)^{-n_kp/2}etr(-\frac{1}{2\sigma_k^2}\mathbf{S}_k)\left(\prod_{i=1}^r(1-\omega_{ki})
   \right) ^{n_k/2}
   etr(\frac{1}{2\sigma_k^2}(VO_k\Omega_kO_k^TV^T)\mathbf{S}_k)
\label{eqn:sslik}
\end{equation}

We use conjugate and semiconjugate priors for the parameters $O_k$,
$\sigma^2_k$ and $\Omega_k$ to facilitate inference.  In the absence
of real prior information, invariance considerations suggest the use
of priors that lead to equivariant estimators.  Below we describe our
choices for the prior distributions of each parameter and the
resultant posterior distributions.

@@

\begin{align}
\label{lik_vo}
 p(O_k | \sigma^2_k, U_k, \Omega_k) & \propto etr(\Omega_kO^T_k\hat{V}^T[S_k/(2\sigma^2_k)]\hat{V}O_k)
\end{align}

\noindent This is a Bingham$(\Omega, \hat{V}^T \bl S_k \hat{V}/(2\sigma^2)$
distribution on $\mathcal V_{s, p}$ \citep{Khatri1977}. A
Gibbs sampler to simulate from this distribution is given in
\citet{Hoff2012}.

@@

\begin{align}
p(\omega_{ki}|V, O_k, \bl S_k) \propto_{\omega_{ki}}
  \left(\prod_{i=1}^r(1-\omega_{ki})^{n_k/2}  \right)
  etr(\frac{1}{2\sigma_k^2}(VO_k\Omega_kO_k^TV^T)\mathbf{S}_k) \\
&  \propto & (1-\omega_{ki})^{n/2} e^{c \omega_s  n/2},    
\label{eqn:wpost}
\end{align}

@@

% \begin{align}
% p(V|\bl O, \bl S, \bl \sigma^2, \bl \Omega) &\propto etr(\sum_{k=1}^K[S_k/(2\sigma^2_k)]V[O_k\Omega_kO^T_k]V^T)\\
% \end{align}

% Although close in appearance, this conditional distribution is not a matrix
% Bingham-von Mises-Fisher distribution.  However, the conditional distribution
% of a column of $V$ given all other columns is a vector Bingham-von
% Mises-Fisher distribution.  To see this, let
% $A_k = S_k/(2\sigma^2_k)$ and $B_k = O_k\Omega_kO^T_k$.  We can write
% the conditional distribution of $V$ given $A_k$ and $B_k$ as
% \begin{align}
% P(V|A_k,B_k) &\propto etr(\sum_{k=1}^KA_kVB_kV^T)\\
% &= etr(\sum_{k=1}^K\sum_{i,j}^sA_kv_iv_j^Tb_{ijk})\\
% &= etr(\sum_{i,j}^s \sum_{k=1}^K v_j^T[b_{ijk}A_k]v_i)
% \end{align}

@@

% \begin{align}
% P(v_i|A_k,B_kV_{-i})  &\propto etr(2(\sum_{j\neq i}^s v_j^T[\sum_{k=1}^K b_{ijk}A_k])v_i+v_i^T[\sum_{k=1}^K b_{ijk}A_k]v_i))\\
% &= etr(Cv_i+v_i^TMv_i)
% \label{eqn:vbmf}
% \end{align}

@@

\section{Simulation Study}

@@

Given estimates of $s$ and $r$, we then proceed with Bayesian
inference of the eigenvectors, eigenvalues, and noise variance using
each of the three inferential models.  When applying the shared
subspace estimator, we use the EM algorithm presented in Section
\ref{sec:em} first, to estimate $VV^T$ and then use Bayesian inference
to estimate the posterior distribution of all remaining quantities
conditional on $V$.  For all covariance estimates, we compute Stein's
loss
$L_S( \Sigma_k , \hat\Sigma_k) = \text{tr}( \Sigma_k^{-1} \hat
\Sigma_k ) - \log |\Sigma_k^{-1} \Sigma_k | - p$
for each group.  Under Stein's loss, the Bayes estimator is the
inverse of the posterior mean of the precision matrix,
$\hat \Sigma_{k} = \Exp{ \Sigma_k^{-1} | \bl S_k}^{-1}$ which we estimate
using MCMC samples.  In Table \ref{table:groupLoss} we report the
average group loss
$L(\Sigma_1, ..., \Sigma_K; \hat\Sigma_1, ..., \hat\Sigma_K ) =
\frac{1}{K} \sum_k L_S( \Sigma_k , \hat\Sigma_k)$
and 95\% intervals (TODO) for the estimates derived using each
inferential model.

Naturally, the estimates with the lowest loss are derived from the
inferential model that matches the data generating model. However, the
shared subspace estimator has small loss under model misspecification
relative to the alternative.  For example, when the data is
generated using the complete pooling model, the shared subspace
estimator has almost four times smaller loss than the no pooling
estimator.  When the data is generated under the no pooling model, the
shared subspace estimator is over an order of magnitude better than
the complete pooling estimator.

Thus, shared subspace estimation can be an appropriate choice when the
amount of similarity between groups is not known a priori.  The shared
subspace model combined with the rank estimation procedure suggested
by \citet{Gavish2014} leads to an estimator which adapts to the amount
of similarity between groups.  Importantly, the shared subspace model
can be viewed as a generalization of the ``no pooling'' spiked
covariance model.  The two are equivalent when we let
the shared subspace span the full space, e.g. $V \in \mathcal{O}(p)$.
Figure \ref{fig:sdimension} depicts the loss for data generated under
the no pooling model and fit with shared subspace estimators with
different dimension.  As the shared subspace dimension increases to
the dimension of the feature space, $s \rightarrow p$, the loss for
the shared subspace estimator quickly coverges to the loss for the
individual spiked covariance estimator.  Note that
span($[U_1, ..., U_K]) \leq rK$, thus even when there is little
similarity between the eigenvectors from each group, the shared
subspace estimator with $s = rK$ will perform comparably to the ``no
pooling'' estimator, provided that we can identify a subspace where
span($\hat{V})$ is close to $\text{span}([U_1, ..., U_K])$.

In many real world examples, we are more interested in finding a
useful space to interpret and compare differences between covariance
matrices than we are in finding the lowest-loss estimator for the full
covariance matrices.  In this context, the relevant quantity is often
an estimate of the projected covariance, $V^T\Sigma V$.  In Figure
\ref{fig:lossBoxplot} we depict boxplots for the distribution of
losses,
$L_S(\hat{V}^T\Sigma_k\hat{V}, \hat{V}^T\hat{\Sigma}_k\hat{V})$, on
simulated data.  In the left boxplot the data come from a
2-dimensional shared subspace model and on the right plot the data
come from a 10-dimensional shared subspace model.  For both boxplots,
estimates
$\hat{V}^T\hat{\Sigma}_k\hat{V}$ are derived from assumed 2-dimensional
model. This figure demonstrates that the loss for estimating the
projected covariance matrix is not significantly larger when we
underestimate the dimension of the true shared subspace.

\begin{table}
\begin{center}
  \begin{tabular}{ l  l | c | c | c |}
    \multicolumn{2}{c}{} & \multicolumn{3}{c}{\textbf{Inferential Model}} \\
  \multicolumn{2}{c|}{}  & Shared Subspace & Complete Pooling & No pooling \\  \cline{2-5}
    \multirow{3}{*}{\rotatebox[origin=c]{90}{\textbf{Data Model}}} & 
   Shared Subspace & 0.8 & 2.1 & 3.0 \\ \cline{2-5}
   & Complete Pooling & 0.8 & 0.7 & 3.0 \\ \cline{2-5}
   & No pooling & 8.3 & 136.3 & 3.0 \\ \cline{2-5}
  \end{tabular}
  \caption[Table caption text]{Stein's loss for different inferential
    and data generating models.  For each of the $K=10$ groups, we
    simulate a $p=200$ dimensional spiked covariance matrix for rank
    $r=2$.  For data from a shared subspace model the eigenvectors
    of the covariance matrix share a subspace dimension $s=2$.  All results are
    based on $n_k = 50$ observations per group.  }
\label{table:groupLoss}
\end{center}
\end{table}

>>>>

\begin{figure}[t]
    \centering
    \begin{subfigure}[b]{0.25\textwidth}
        \includegraphics[width=\textwidth]{Figs/3dplot}
        \caption{Ratio}
        \label{fig:dmelanRatio}
    \end{subfigure}
    \qquad     \qquad     \qquad 
 %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
      %(or a blank line to force the subfigure onto a new line)
    \begin{subfigure}[b]{0.25\textwidth}r
        \includegraphics[width=\textwidth]{Figs/2d-scatter}
        \caption{Posterior}
        \label{fig:dmelanPosterior}
    \end{subfigure}
    \caption{ }
\end{figure}

\section{Inference}

In this section we outline Bayesian inference algorithms including
both an optimization algorithm and an MCMC algorithm for estimating
covariance parameters and their associated uncertainty.  MCMC
inference is particularly appealing because of the ease with which we
can quantify the uncertainty associated with each estimate.
Uncertainty quantification is particularly important in the $n << p$
setting where seemingly large differences in the point estimates of
the low rank covariance matrices across groups may not actually
reflect significant differences.  Although, uncertainty quantification
for all parameters is desirable, for the size of the problems we are
generally interested in, we found MCMC inference for the subspace
$VV^T$ to be intractable.  Our optimziation based approached is
signifcantly faster and works well for large large dimensional
datasets.  As such, we start, in Section \ref{sec:em}, by outlining an
expectation-maximization algorithm for estimating the shared subspace
and derving point estimates of the projected covariance $\psi_k$. In
Section \ref{sec:bayes} we describe a Gibbs sampler for posterior
inference of $\psi_k$ given the common subspace.  These two algorithms
can be used in conjuction to first identify and evaluate a shared
subspace and then to quantify the uncertainty associated with the
projected covariance matrices on this space.

\subsection{An Expectation-Maximization Algorithm}
\label{sec:em}

In this section we describe a computationally efficient algorithm for
estimating the shared subspace, $VV^T$, projected covariance matrix
$\psi_k$ and noise parameter $\sigma_k^2$.  First, note that the full
likelihood for the shared subspace model can be written as

\begin{align}
p(S_1, ... S_k | \Sigma_k,n_k) &\propto \prod_{k=1}^K |\Sigma_k|^{-n_k/2}\etr(-\Sigma_k^{-1}\mathbf{S}_k/2)  \\
&\propto \prod_{k=1}^K  |\Sigma_k|^{-n_k/2}\etr(-(\sigma_k^2(V\psi_kV^T +
  I))^{-1}\mathbf{S}_k/2) \\
&\propto \prod_{k=1}^K  |\Sigma_k|^{-n_k/2}\etr(-\left[V(\psi_k +
  I)^{-1}/\sigma_k^2 V^T + (I-VV^T)/\sigma^2_k\right]\mathbf{S}_k/2)
  \\
&\propto \prod_{k=1}^K  (\sigma_k^2)^{-n_k(p-s)/2}|M_k|^{-n_k/2}\etr(-\left[VM_k^{-1}V^T + \frac{1}{\sigma^2_k} (I-VV^T)\right]\mathbf{S}_k/2) 
\end{align}

@@

% \begin{align}
% p(V|...) &\propto \etr(\sum_{k=1}^K[O_k\Omega_kO^T_k]V^TS_k/(2\sigma^2_k)]V)\\
% \end{align}

We propose an expectation-maximization (EM)
algorithm, to estimate the relevant quantities.  In particular we
maximize the marginal posterior distribution of $V$, treating 
$M_k^{-1}$ and $\frac{1}{\sigma_k^2}$ as the ``missing'' parameters to
be imputed.  We assume independent Jeffreys' prior distributions for
both $\frac{1}{\sigma_k^2}$ and $M_k^{-1}$ and a uniform prior for $V
\in \mathcal{V}_{s, p}$.  In particular we let
$p(\frac{1}{\sigma_k^2}) \propto \sigma_k^2$ and
$p(M_k^{-1}) \propto |M_k^{-1}|^{-(s+1)/2}$.  From the likelihood it
can easily be shown that the conditional posterior for $M_k^{-1}$ is

$$p(M_k^{-1} | V) \propto |M_k^{-1}|^{(n_k - s -1)/2}\etr(-1/2(M_k^{-1}V^T\mathbf{S}_kV)) $$

@@

$$p\left(\frac{1}{\sigma^2} | V\right) \propto \left(\frac{1}{\sigma_k^2}\right)^{n_k(p-s)/2-1}\etr\left(-\frac{1}{2\sigma^2_k} (I-VV^T)\mathbf{S}_k\right)  $$

@@

\paragraph{Evaluating a Shared Subspace}

The above EM algorithm provides an estimate of a shared subspace, but
in practice there are no guarantees that the inferred subspace is
equally appropriate for all groups.  Moreover, the shared subspace can
be chosen in other ways, including more ad hoc methods
involving the singular value dcomposition or simply by incorporating
subject matter considerations.  Regardless of how a particular
subspace is identified, it is usually important to quantify how useful
this subspace is for explaining the variability in each group.  In
particular, if $\hat{V}$ is a basis for a subspace shared across
groups, then for each $k$, most of the variation in $\Sigma_k$ should
lie on the subspace spanned by $\hat{V}$.  To quantify the extent to
which this is true for different groups, we propose a simple estimator
for the proportion of ``signal'' variance that lies on any given
subspace.  Specifically, we apply the following estimator for ratio of
the sum of the first $s$ eigenvalues of
$\hat{V}^T\bl \Sigma_k \hat{V}$ to the sum of the first $s$
eigenvalues of $\Sigma_k$:

where $||.||_F$ is the Frobenius norm. If the inferred subspace is the
true subspace, e.g. $\hat{V}\hat{V}^T = VV^T$, then this estimator
will asymptotically approach one.  To see this, note that
$\underset{V \in \mathcal{V}_{p, s}}{sup}\frac{||Y_k^TV||_F}{n_k}$ is
equivalent to the sum of the first $s$ eigenvalues of the sample
covariance matrix $\frac{\bl S_k}{n_k}$.  With $\hat{\lambda}^{(k)}_i$
the $i$-th eigenvalue of $\frac{\bl S_k}{n_k}$, \citet{Baik2006}
demonstrated that asymptotically as $p, n_k \rightarrow \infty$ with
$\frac{p}{n_k} = \gamma_k$ fixed

Thus,
$\sum_i^s \hat{\lambda}^{(k)}_i \approx \sum_i^s \lambda^{(k)}_i +
\sigma_k^2ps$.
If we assume that $s$ is fixed in this asymptotic regime, then
$||Y_k^TV||_F/n_k$ is a consistent estimator for
$\sum_i^s \lambda^{(k)}_i$, and thus the above estimator will be close
to one when $\hat{V}\hat{V}^T = VV^T$ and smaller if not.  The metric provides a
useful indicator for which groups can be reasonably compared on a
given subspace and which groups cannot.  In Section \ref{sec:simulation} we plot
these estimates for different simulated datasets.  In Section \ref{sec:bio}
we plot the estimates for inferred subspaces for real biological data.

\subsection{Uncertainty Quantification}
\label{sec:bayes}

Although the EM algorithm presented in the previous section yields a
fast way to derive point estimates for $VV^T$ and $\psi_k$, it does
not lead to natural uncertainty bounds for these estimates. Full
uncertainty quantification fo $VV^T$ is desirable but we found this
challenging for the dimensions of the problems we would like to
consider.  As such, in this section, we assume that the subspace
$VV^T$ is fixed and known and focus on uncertainty quantificatoin for
$\psi_k$ only.  Further, by focusing on group differences on a fixed
subspace, it is simpler to interpret differences across covariance
matrices.  The shared subspace can be an estimated using the EM algorithm
presented in Section \ref{sec:em} or chosen based on subejct matter
considerations.  Note that when the subspace is known, the posterior
distribution of $\Sigma_k$ is conditionally independent from the other
groups, i.e.
$P(\Sigma_{k} | V, \bl S_k, \Sigma_{-k}) = P(\Sigma_{k} | V, \bl
S_k)$,
so that we can independently estimate the conditional posterior
distribution for each group.

There are many different prior specifications for $\psi_k$ and
$\sigma_k^2$ that we could consider for estimating posterior
uncertainty.  One simple approach would be to use the same prior
specification used in Section \ref{sec:em}.  That is, we can use
Jeffreys' prior for $\frac{1}{\sigma_k^2}$ and
$M_k^{-1} = (\psi_k + I) ^{-1}/\sigma_k^2$.  In fact, it is easy to
show that the Gamma distribution for $\frac{1}{\sigma_k^2}$ and the
Wishart distribution for $M_k^{-1}$ are conjugate priors for the
respective quantities.  From posterior samples of $M_k$ and
$\sigma_k^2$ we can easily recover samples of $\psi_k$.
Unfortunately, if $M_k^{-1}$ has a Wishart distribution, the implied
support of $\psi_k$ is not constrained to the cone of positive
semi-definite matrices.

Direct prior specifications for $\psi_k$ are more appealing, but many
obvious choices, like the inverse-Wishart prior distribution for
$\psi_k$ do not lead to simple posterior distributions.  As such,
building on recent interest in spiked covariance models
\citep{Hoff2009, Donoho2013, Paul2007} we focus on tractable models
based on the eigendecomposition of $\psi_k$.  Such models are
especially appealing because it is simple to summarize and compare
$\psi_k$ across groups in terms their eigenstructure (See Section
\ref{sec:simulation} and \ref{sec:bio}).

By modeling the eigenstructure, we can now view each covariance
$\Sigma_k$ as a spiked covariance matrix.  Equation \ref{eqn:sspsi}, written as a function $V$, becomes

@@

Here, we make one further generalization to allow
$\psi_k = O_k\Lambda_KO_k^T$ be a rank $r \leq s$ dimensional covariance
matrix on the $s$-dimensional subspace.  Thus, $\Lambda_k$ is an $r \times
r$ diagonal matrix of eigenvalues, and $O_k \in
\mathcal{V}_{s,r}$ is the matrix of eigenvectors of
$\psi_k$.  For any individual group, this corresponds to the original spiked
covariance model (Equation \ref{eqn:spiked}) with $U_k = VO_k \in
\mathcal{V}_{p, r}$.

@@

\begin{equation}
p(S | \Sigma_k,n_k) \propto |\Sigma_k|^{-n/2}\etr(-\Sigma_k^{-1}\mathbf{S}_k/2)
\end{equation}

@@

Where the diagonal matrix $\Omega = \Lambda(I+\Lambda)^{-1}$, e.g. $\omega_i = \frac{\lambda_i}{\lambda_{i}+1}$.  Further,

@@

where the second line is due to Sylvester's determinant theorem.  The
likelihood of $V$, $O_k$, $\Lambda_k$ and $\sigma_k^2$ is available
from Equation\ref{eqn:rplik} by substituting in the appropriate
quantities in for $\Sigma^{-1}_k$ and $|\Sigma_k|$ and replacing $U_k$
with $VO_k$:

\begin{equation}
 L(\sigma_k^2,\bl V , O_k \Omega_k : \bl Y_k) \propto
    (\sigma_k^2)^{-n_kp/2}\etr(-\frac{1}{2\sigma_k^2}\mathbf{S}_k)\left(\prod_{i=1}^r(1-\omega_{ki})
   \right) ^{n_k/2}
   \etr(\frac{1}{2\sigma_k^2}(VO_k\Omega_kO_k^TV^T)\mathbf{S}_k)
\label{eqn:sslik}
\end{equation}

We use conjugate and semiconjugate priors for the parameters $O_k$,
$\sigma^2_k$ and $\Omega_k$ to faciliate inference via a Gibbs
sampling algorithm.  In the absence of real prior information,
invariance considerations suggest the use of priors that lead to
equivariant estimators.  Below we describe our choices for the prior
distributions of each parameter and the resultant conditional posterior
distributions.

@@

\begin{align}
\label{lik_vo}
 p(O_k | \sigma^2_k, U_k, \Omega_k) & \propto \etr(\Omega_kO^T_k\hat{V}^T[S_k/(2\sigma^2_k)]\hat{V}O_k)
\end{align}

\noindent This is a Bingham$(\Omega, \hat{V}^T \bl S_k \hat{V}/(2\sigma^2)$
distribution on $\mathcal V_{s, r}$ \citep{Khatri1977}. A
Gibbs sampler to simulate from this distribution is given in
\citet{Hoff2012}.

@@

\begin{align}
p(\omega_{ki}|V, O_k, \bl S_k) \propto_{\omega_{ki}}
  \left(\prod_{i=1}^r(1-\omega_{ki})^{n_k/2}  \right)
  \etr(\frac{1}{2\sigma_k^2}(VO_k\Omega_kO_k^TV^T)\mathbf{S}_k) \\
&  \propto & (1-\omega_{ki})^{n/2} e^{c \omega_s  n/2},    
\label{eqn:wpost}
\end{align}

@@

% \begin{align}
% p(V|\bl O, \bl S, \bl \sigma^2, \bl \Omega) &\propto \etr(\sum_{k=1}^K[S_k/(2\sigma^2_k)]V[O_k\Omega_kO^T_k]V^T)\\
% \end{align}

% Although close in appearance, this conditional distribution is not a matrix
% Bingham-von Mises-Fisher distribution.  However, the conditional distribution
% of a column of $V$ given all other columns is a vector Bingham-von
% Mises-Fisher distribution.  To see this, let
% $A_k = S_k/(2\sigma^2_k)$ and $B_k = O_k\Omega_kO^T_k$.  We can write
% the conditional distribution of $V$ given $A_k$ and $B_k$ as
% \begin{align}
% P(V|A_k,B_k) &\propto \etr(\sum_{k=1}^KA_kVB_kV^T)\\
% &= \etr(\sum_{k=1}^K\sum_{i,j}^sA_kv_iv_j^Tb_{ijk})\\
% &= \etr(\sum_{i,j}^s \sum_{k=1}^K v_j^T[b_{ijk}A_k]v_i)
% \end{align}

@@

% \begin{align}
% P(v_i|A_k,B_kV_{-i})  &\propto \etr(2(\sum_{j\neq i}^s v_j^T[\sum_{k=1}^K b_{ijk}A_k])v_i+v_i^T[\sum_{k=1}^K b_{ijk}A_k]v_i))\\
% &= \etr(Cv_i+v_i^TMv_i)
% \label{eqn:vbmf}
% \end{align}

@@

\section{Simulation Study}
\label{sec:simulation}

\subsection{Model Works}
First, we do a single simulation that shows that we can recover the
true parameters using our inferential algorithm when the assumed model
is correct.

\begin{figure}[t]
    \centering
    \begin{subfigure}[b]{0.45\textwidth}
    \includegraphics[width=\textwidth]{Figs/posteriorRegions}
        \caption{Posterior}
        \label{fig:simPosterior}
    \end{subfigure}
    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
      %(or a blank line to force the subfigure onto a new line)
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{Figs/simRatio-05-19}
        \caption{Ratio}
        \label{fig:simRatio}
    \end{subfigure}
    \caption{a) 95\% posterior regions for the ratio of the eigenvalues,
      $\frac{\lambda_1}{\lambda_2}$, of $\psi_k$ and the orientation
      of the principle axis on space spanned by $\hat{V}$ cover the
      truth.  True parameter values on $\hat{V}^T\Sigma_k\hat{V}$ are
      indicated by the crosses.  Since $V$ is only identifiable up to
      rotation we find the Procrustes rotation that maximizes the
      similarity of $\hat{V}$ to the true data generating basis. True
      eigenvalue ratios were 10 (red and black), 3 (green and blue)
      and 1 (light blue).  True orientations were $\pi/4$ (black), $-\pi/4$
      (red and green) and 0 (blue and cyan).  b)  Estimated proportion
    of variance explained by the subspace (Equation \ref{}).  As expected, for each
    group, all of the variance can be explained on the inferred subspace.  }
\end{figure}

In order to quantify the uncertainty associated with the relevant
parameters conditional on a shared subspace, we use the Bayesian
inference algorithm outlined in Section \ref{sec:bayes}. In Figure
\ref{fig:posteriorRegions} we depict the 95\% posterior regions for
statistics based on the eigendecomposition of
$\psi_k$.  Specifically, for the two dimensional
shared subspace model, we plot the eigenvalue ratio
$\frac{\lambda_1}{\lambda_2}$ and the angle of the first eigenvector
on the $\text{arctan}(\frac{O_2}{O_1})$, where
$\lambda_1 > \lambda_2$.  Non-overlapping 95\% posterior regions are
indicative of significant differences in the covariance structure
between groups.

In this example, the posterior regions for all groups cover the true
eigenvalue ratio and orienntation of the primary axis for the matrix
$\hat{V}^T\Sigma_k\hat{V}$.  Note that for smaller eigenvalue ratios,
there is larger uncertainty about the orientation of the primary axis.
When the ratio is one (light blue) there is no information about
orientation of the primary axis since the contours are spherical.

\subsection{Model Misspecification}

@@

Given estimates of $s$ and $r$, we then proceed with Bayesian
inference of the eigenvectors, eigenvalues, and noise variance using
each of the three inferential models.  When applying the shared
subspace estimator, we use the EM algorithm presented in Section
\ref{sec:em} first, to estimate $VV^T$ and then use Bayesian inference
to estimate the posterior distribution of all remaining quantities
conditional on $V$.  For all covariance estimates, we compute Stein's
loss
$L_S( \Sigma_k , \hat\Sigma_k) = \text{tr}( \Sigma_k^{-1} \hat
\Sigma_k ) - \log |\Sigma_k^{-1} \Sigma_k | - p$
for each group.  Under Stein's loss, the Bayes estimator is the
inverse of the posterior mean of the precision matrix,
$\hat \Sigma_{k} = \Exp{ \Sigma_k^{-1} | \bl S_k}^{-1}$ which we
estimate using MCMC samples.  In Table \ref{table:groupLoss} we report
the average group loss
$L(\Sigma_1, ..., \Sigma_K; \hat\Sigma_1, ..., \hat\Sigma_K ) =
\frac{1}{K} \sum_k L_S( \Sigma_k , \hat\Sigma_k)$
and 95\% intervals for the estimates derived using each inferential
model.

Naturally, the estimates with the lowest loss are derived from the
inferential model that matches the data generating model. However, the
shared subspace estimator has small loss under model misspecification
relative to the alternative.  For example, when the data is generated
using the complete pooling model, the shared subspace estimator has
almost four times smaller loss than the no pooling estimator.  When
the data is generated under the no pooling model, the shared subspace
estimator is over an order of magnitude better than the complete
pooling estimator.  Thus, shared subspace estimation can be an
appropriate choice when the amount of similarity between groups is not
known a priori.

The shared subspace model combined with the rank
estimation procedure suggested by \citet{Gavish2014} leads to an
estimator which adapts to the amount of similarity between groups.
Importantly, the shared subspace model can be viewed as a
generalization of the ``no pooling'' spiked covariance model.  The two
are equivalent when we let the shared subspace span the full space,
e.g. $V \in \mathcal{O}(p)$.  Figure \ref{fig:sdimension} depicts the
loss for data generated under the no pooling model and fit with shared
subspace estimators with different dimension.  As the shared subspace
dimension increases to the dimension of the feature space,
$s \rightarrow p$, the loss for the shared subspace estimator quickly
coverges to the loss for the individual spiked covariance estimator.
Note that span($[U_1, ..., U_K]) \leq rK$, thus even when there is
little similarity between the eigenvectors from each group, the shared
subspace estimator with $s = rK$ will perform comparably to the ``no
pooling'' estimator, provided that we can identify a subspace where
span($\hat{V})$ is close to $\text{span}([U_1, ..., U_K])$.

For any given subspace we can identify whether the space is
inappropriate for some groups, either because the dimension is too
small or the space itself does not represent every group equally.
Specifically, we estimate which groups are well represented on the
subspace and which groups are not using the proportion of variance
estimator from Section \ref{sec:em} (Equation \label{eqn:ratio}).
Figure \ref{fig:evalRatios} depicts the estimated proportion of
variance on the subspace for each group in two different simulated
datasets.  When the shared subspace model is appropriate, the subspace
spanned by $\hat{V}$ explains most of the variance in $\Sigma_k$, for
all groups.  When the estimated subspace is too small, the proportion
of variance will be small for at least some of the groups.

Ultimately, in many real world examples, we are more interested in
finding a useful space to interpret and compare differences between
covariance matrices than we are in finding the lowest-loss estimator
for the full covariance matrices.  In this context, the relevant
quantity is often an estimate of the projected covariance,
$V^T\Sigma V$.  In Figure \ref{fig:lossBoxplot} we depict boxplots for
the distribution of losses,
$L_S(\hat{V}^T\Sigma_k\hat{V}, \hat{V}^T\hat{\Sigma}_k\hat{V})$, on
simulated data.  In the left boxplot the data come from a
2-dimensional shared subspace model and on the right plot the data
come from a 10-dimensional shared subspace model.  For both boxplots,
estimates $\hat{V}^T\hat{\Sigma}_k\hat{V}$ are derived from assumed
2-dimensional model. This figure demonstrates that the loss for
estimating the projected covariance matrix is not significantly larger
when we underestimate the dimension of the true shared subspace.
